Article Title

### 2014-11-21 manuscript.tex
minor

Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed nearly equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances. 
Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
With increasing sample sizes, all estimators converged to the truth (zero loss).  
However, the discriminability between their performances only increased with sample size (data not shown).

In more realistic conditions, when the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
Then the \emph{validation loss} function $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 1 p$ and omitting the constant term: 
\begin{equation}\label{eq:vloss}
    \loss{C,C^\prime_{\sf sample}} = \frac 1 {2p} \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
\end{equation}

Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, validation loss is an unbiased estimate of true loss up to a constant:
\begin{equation}
    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
\end{equation}
Therefore, the validation procedure allows comparing the relative values of loss of different covariance estimators.

Indeed, validation loss computed by 10-fold cross-validation (see Methods) accurately reproduces the relative values of the loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).

In the example above, the data were sampled from  multivariate normal random variables. In such models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data. 
To demonstrate that estimator rankings were robust to deviations from Gaussian models, as long as dependencies between variables included a strong linear component, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.

@@

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}). The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

@@

We evaluated the covariance matrix estimators a loss function based on multivariate normal log likelihood.  However, this does not limit the applicability of its conclusions to normal distributions. Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.  If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.

@@

When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.  This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:5}--\ref{fig:7}).

>>>>

Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly inferred zero factors and only estimated the individual variances. 
Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed nearly equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
With increasing sample sizes, all estimators converged to the truth (zero loss).  
However, the discriminability between their performances only increased with sample size (data not shown).

In more realistic conditions, when the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
Then the \emph{validation loss} $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 1 p$ and omitting the constant term: 
\begin{equation}\label{eq:vloss}
    \loss{C,C^\prime_{\sf sample}} = \frac 1 {2p} \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
\end{equation}

Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, the validation loss is an unbiased estimate of the true loss, up to a constant:
\begin{equation}
    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
\end{equation}
Therefore, the validation procedure allows comparing the relative values of loss of different covariance estimators.

Indeed, the validation loss computed by 10-fold cross-validation (see Methods) accurately reproduces the relative values of the true loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).

In the example above, the data were sampled from  multivariate normal random variables. In such models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data. 
To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.

@@

To test this hypothesis, we computed the relative validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}). The respective median differences of the validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

@@

We evaluated the covariance matrix estimators using a loss function derived from the normal distribution.  However, this does not limit the applicability of its conclusions to normal distributions. Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.  If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.

@@

When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.  This approach was used to compute the covariance matrix estimates and their true loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:5}--\ref{fig:7}).
